---
title: "Exponential Distribution in R"
author: "Rehab Fathi"
date: "April 6, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Overview
In this project, I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. I will investigate the distribution of averages of 40 exponentials with a thousand simulations.

## Simulations
```{r}
par(mfrow=c(1,2))
hist(rexp(40,0.2), main="Exponential Distribution", xlab="Exponentials")
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40,0.2)))
hist(mns, main = "Distribution of the Mean", xlab = "Mean of 40 Exponentials")
```

The figures show the exponential distribution compared to the means of 1000 simulations of 40 samples of the exponential distribution. The rate parameter is set to 0.2



## Sample mean versus theoretical mean
```{r}
par(mfrow=c(1,2))
hist(rexp(40,0.2), main="Exponential Distribution", xlab="Exponentials")
abline(v=mean(rexp(40,0.2)), lty =1, lwd = 5)
hist(mns, main = "Distribution of the Mean", xlab = "Mean of 40 Exponentials")
abline(v=mean(mns), lty =1, lwd = 5)
```

The mean of the sample means is 5.41, slightly higher than the exponential mean which is 5.0



## Sample variance versus theoretical variance
```{r}
par(mfrow=c(1,2))
hist(rexp(40,0.2), 
     main="Exponential Distribution", 
     xlab="Exponentials", 
     prob=TRUE)
curve(dnorm(x, mean=mean(rexp(40,0.2)), sd=sqrt(var(rexp(40,0.2)))),
          col="darkblue", lwd=2, add=TRUE, yaxt="n")
abline(v=mean(rexp(40,0.2)) + sqrt(var(rexp(40,0.2))), lwd=2, col="red")
abline(v=mean(rexp(40,0.2)) - sqrt(var(rexp(40,0.2))), lwd=2, col="red")
hist(mns, 
     main = "Distribution of the Mean", 
     xlab = "Mean of 40 Exponentials", 
     prob=TRUE)
abline(v=mean(mns) + sqrt(var(mns)), lwd=2, col="red")
abline(v=mean(mns) - sqrt(var(mns)), lwd=2, col="red")
curve(dnorm(x, mean=mean(mns), sd=sqrt(var(mns))), 
          col="darkblue", lwd=2, add=TRUE, yaxt="n")
```

The distribution of means has a lower variance.
The variance of the sample of means is 0.67 while the theoretical variance of the exponential distribution is 2.23



## Sample distribution
```{r}
hist(mns, 
     main = "Distribution of the Mean", 
     xlab = "Mean of 40 Exponentials", 
     prob=TRUE)
curve(dnorm(x, mean=mean(mns), sd=sqrt(var(mns))), 
          col="darkblue", lwd=2, add=TRUE, yaxt="n")
lines(density(mns), lwd=2, col="red")   
```

The red curve shows the distribution of the sample of means while the blue curve shows the normal distribution. The distribution of means is a little skewed but it is approximately a normal distribution.
Another property of a normal distribution is that the mean is equal to the median, the mean of the distribution is 5 and the median is 4.93.