---
title: "ToothGrowth Data"
author: "Rehab Fathi"
date: "April 6, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## OVerview

The dataset shows the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

## Summary of the data
```{r}
df <- ToothGrowth
summary(df)
```

## Data Analysis
Let's explore the average length for each dose of vitamin C for the two sources
```{r}
summ <- aggregate(df$len,
          by=list(df$supp, df$dose),
          FUN="mean")
names(summ) <- c("Source", "Dose", "Mean")
summ
```

## Hypothesis
My hypothesis is that the mean odontoblasts length with orange juice is equal to the mean length with ascorbic acid.

```{r}
t.test(summ[summ$Source=="OJ","Mean"]- summ[summ$Source=="VC","Mean"])
```

## Conclusion
The null hypothesis is rejected by the t-test. The mean length associated with orange juice is not equal to the mean length associated with ascorbic acid.
