---
title: "Regression Models"
author: "Rehab Fathi"
date: "April 11, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Executive Summary

I work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

- "Is an automatic or manual transmission better for MPG"
- 
"Quantify the MPG difference between automatic and manual transmissions"

In the simple model, the MPG difference is 7.245 MPG, and in the multivariate enhanced model, the MPG difference is 1.47.

## Data Analysis

```{r}
data(mtcars)
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("automatic", "manual")
par(mfrow = c(1,2))
hist (mtcars$mpg[mtcars$am == "manual"],
      xlab = "Manual",
      main= "Miles / gallon",
      col = "red")
hist (mtcars$mpg[mtcars$am == "automatic"],
      xlab = "Automatic",
      main= "Miles / gallon",
      col = "red")


```


```{r}
agg <- aggregate(mtcars$mpg, 
          by = list(mtcars$am),
          mean)
names(agg) <- c("Type", "Mean mpg")
agg
```

## Hypothesis

Automatic cars have lower mpg than manual cars

## T-Test

```{r}
manual <- mtcars[mtcars$am=="manual",]
automatic <- mtcars[mtcars$am=="automatic",]
t.test(automatic$mpg, manual$mpg)
```

The p-value is 0.001374 showing a significant difference

## Regression Model

```{r}
summary(lm(mtcars$mpg ~ mtcars$am))
```

The average mpg for manual cars is 7.245 higher than automatic cars. The model explains 36% of the variance in the data

## Multivariate Regression

```{r}
summary(lm(formula = mpg ~ cyl + hp + wt + am, data = mtcars))
```

This model explains 85% of the variance
Manual cars have higher mpg by 1.47


## Plot Residuals

```{r}
par(mfrow = c(2,2))
plot(lm(formula = mpg ~ cyl + hp + wt + am, data = mtcars))
```

